---
title: "Homework 1"
author: "PSTAT 134/234"
format: html
editor: visual
---

## Homework 1

For this assignment, you may use R or Python, or a combination of both, to complete both case studies. You can use the code that we include in Labs 1 and 2 to answer these questions. You also may need to use other functions. I encourage you to make use of our textbook(s) and use the Internet to help you solve these problems. You can also work together with your classmates. If you do work together, you should provide the names of those classmates below.

[Names of Collaborators (if any): Anthony Cu]{.underline}

### Case Study: New York Times Ad Impressions (Simulated)

There are 10 data sets in the `/data` subdirectory named `nyt1.csv`, `nyt2.csv`, ..., `nyt10.csv`. Each file represents one day's worth of simulated data on ad impressions and clicks on the [New York Times homepage](https://www.nytimes.com/). Each row represents a single user. There are five columns:

-   `Age` (user's age)

-   `Gender` (user's gender, coded as `0` = female, `1` = male)

-   `Impressions` (number of ads displayed during the user's visit)

-   `Clicks` (number of clicks made by the user)

-   `Signed_In` (whether or not the user was signed in as a member)

Load **a single data file**. Then do the following.

1.  Create a new variable, `age_group`, that categorizes users into the following age groups: \< 18, 18-24, 25-34, 35-44, 45-54, 55-64, and 65+.

2.  Plot the distributions of impressions and "click-through rate" for all 6 age categories. (*Note:* Click-through rate is defined as the number of clicks divided by the number of impressions; it represents the proportion of ads that generated clicks.)

3.  Create a new variable to categorize users based on their click behavior. (The name and categories for this variable are up to you. Explain what decision\[s\] you make and why.)

4.  Explore the data and make visual and quantitative comparisons across user segments/demographics to answer the following:

    -   How do \<18 year old males differ from \<18 year old females in terms of clicks and impressions?

    -   How does the distribution of click-through rate for users who are signed in differ from the distribution for those who are **not** signed in?

    -   Are certain age groups more likely to be signed in than others? Which ones?

5.  Calculate summary statistics for the click-through rate. These should include (1) quartiles, (2) mean, (3) median, (4) minimum and maximum, and (5) variance. Choose two user segments to compare these statistics across (for example, compare the mean, median, and quartiles for users 25-34 to those for users 65+).

6.  Summarize your findings in a brief (1-2 paragraph) report intended for a New York Times (NYT) advertising team.

```{r}
library(tidyverse)
library(ggplot2)
```

```{r}
# read in data
ny_data <- read_csv("~/pstat-134-234/Homework/Homework 1/data/nyt6.csv")
```

### 1

```{r}
ny_data$age_group <- cut(ny_data$Age,
                         breaks = c(0,18,25,35,45,55,65,112),
                         labels = c("<18","18-24","25-34","35-44","45-54","55-64","65+"),
                         include.lowest = T)
ny_data$age_group <- factor(ny_data$age_group)
```

### 2

```{r}
# plotting impressions for each age group
ggplot(ny_data, aes(x = Impressions, fill = age_group)) +
  geom_histogram(alpha = 0.5, binwidth = 1) + 
  labs(title = "Impressions by Age Group", x = "Impressions", y = "Count") +
  theme_minimal()
```

```{r}
# creating click_through_rate variable
ny_data$click_through_rate <- ny_data$Clicks / ny_data$Impressions 
ny_data$click_through_rate[is.na(ny_data$click_through_rate)] <- 0
```

```{r}
# plotting density of CTE for age groups
ggplot(ny_data, aes(x = click_through_rate, fill = age_group)) +
  geom_density(alpha = 0.5) + 
  labs(title = "CTR for each age group", x = "Click through rate", y = 'Count') +
  xlim(c(0, 0.3)) +
  theme_minimal()
```

### 3

```{r}
# creating new variable for impressions + clicks
ny_data$Interactions <- ny_data$Impressions + ny_data$Clicks

# investigating distribution of new variable
table(ny_data$Interactions)
summary(ny_data$Interactions)

# creating categorical variable based on levels of activity
breaks <- c(-1,4,8,Inf)
labels <- c("Low","Medium","High")
ny_data$Activity <- cut(ny_data$Interactions,
                        breaks = breaks,
                        labels = labels)
```

I wanted my new variable to capture the overall activity of the observations while also splitting the observations into groups based on the levels of activity. Summing impressions and clicks gives a full amount of interactions for a particular observation while dividing the results into a factor will help show the general activity levels of the different age groups.

### 4
```{r}
# difference between men and women under 18
ny_data$Gender <- factor(ny_data$Gender,
                         levels = 0:1,
                         labels = c("Female","Male"))

ny_data %>% 
  filter(age_group == '<18') %>%
  select(Gender,Impressions,Clicks,Activity) %>%
  ggplot(aes(x = Impressions, fill = Gender)) + 
  geom_histogram(binwidth = 1) +
  labs(title = "Impressions for Users 18 & Under")

ny_data %>% 
  filter(age_group == '<18') %>%
  select(Gender,Impressions,Clicks,Activity) %>%
  ggplot(aes(x = Clicks, fill = Gender)) + 
  geom_histogram(binwidth = 1) +
  labs(title = "Clicks for Users 18 & Under")
```
  
The plots show that the distribution for impressions is very similar shape but females have significantly more observations than males. This result for the data could suggest that females are more likely to be looking at ads in the New York Times. The distribution for clicks shows that the distributions are similar again, except females have significantly more observations of clicks larger than two. This result could suggest that not only are females viewing NYT ads than men, but also engaging in them more. 

```{r}
# distribution of click through rate for those signed in and not
ny_data$Signed_In <- factor(ny_data$Signed_In,
                            levels = 0:1,
                            labels = c("Not Signed In","Signed In"))

ny_data %>%
  select(Signed_In,click_through_rate) %>%
  ggplot(aes(x = click_through_rate, fill = Signed_In)) +
  geom_density() +
  xlim(c(0,0.2)) +
  labs(title = "Click Through Rate for Signed In & Non-signed in users")
```
  
Most of the click through rates in the dataset are small so it's helpful to look at the rates closer to 0 (even though we're ignoring larger observations). Looking at the density plot for the rates, it appears that those who aren't signed in have slightly higher click through rates than users who are signed in. 

```{r}
# investigating rate of signed in for age groups
ny_data %>%
  select(age_group,Signed_In) %>%
  table()
```
 
According to the table, the only age group that isn't likely to be signed in is users under the age of 18. The rest of the data showed that all other age groups are likely to be signed in, which would suggest adults want to be signed in just in case they act on an advertisement. 
  
### 5

```{r}
tapply(ny_data$click_through_rate, ny_data$age_group, summary)

tapply(ny_data$click_through_rate, ny_data$age_group, var)
```

The average click through rate for users above the age of 65 is about two and a half times larger than the average click through rate for users between 25 and 34. Additionally, the variance of the click through rate for users 65 and older is about three times larger than that of 25-34 year olds. 

### 6

Analyzing the data of impressions on New York Times ads shows demographics where advertisements need to increase clicks and overall engagement We first analyzed impressions across age groups to discover that most individuals see between three and seven advertisements a day. Our analysis also shows that total impressions decreases for older age groups. Additionally, when analyzing the click through rate of ads (impressions divided by the number of clicks), we learned that younger age groups have slightly higher click through rates than older groups. The combination of these analyses reveal to us that advertisements for older demographics need to be better tailored to their interests to increase overall activity and push a higher click through rate. 

Analyzing activity based on gender and other qualities other areas for improvement. Comparing the activity between genders in younger users reveals that younger women (under the age of 18) have a slightly higher activity and click through rate than their male counterparts. Additionally, analyzing whether engagement was when someone was signed in or not revealed that users under 18 are most likely to look at ads without being signed in. These two results suggest we need to push more advertisements catered to young men to encourage them to create an account when engaging in advertisements.

#### Additional Questions for 231 Students

*Now read in at least **three to four more** of these data files and extend your analyses.*

7.  Visualize impressions and click-through rate for signed-in versus signed-out users over time.
8.  Calculate summary statistics to compare signed-in versus signed-out users over time.
9.  Visualize click-through rate for the six different age groups over time.

### Case Study: Social Media Engagement (Simulated)

The data file `Time-Wasters on Social Media.csv` contains a considerable amount of simulated data intended to mimic real-world social media usage scenarios. It comes from this source on Kaggle: <https://www.kaggle.com/datasets/zeesolver/dark-web>

Read through and familiarize yourself with the variables in the dataset. Then answer the following.

10. Produce a summary of the user data (the information about users: age, gender, location, debt, whether they own property, their profession). If you were asked to describe the "average user," what would you say?
11. What video categories are more popular with younger users (up to or below age 20)? What categories are more popular with older users (age 50 or above)? Create a plot or table of the distribution of video categories preferred by younger vs. older users.
12. What platforms are more popular with younger users (up to or below age 20)? What platforms are more popular with older users (age 50 or above)? Create a plot or table of the distribution of platforms preferred by younger vs. older users.

```{r}
# reading in data
social_media_data <- read_csv("~/pstat-134-234/Homework/Homework 1/data/Time-Wasters on Social Media.csv")
```

### 10

```{r}
social_media_data %>%
  select(Age, Debt, `Owns Property`) %>%
  summary()

social_media_data$Profession %>% table()
social_media_data$Gender %>% table()
social_media_data$Location %>% table()
```

I would describe the average user as a 41 year old male who is mostly likely from either India or the US. This average user owns property but is in debt. Students generally aren't 41 years old, so this average user is most likely in labor.

### 11

```{r}
breaks <- c(0,21,50,Inf)
labels <- c("<21","21-49","50+")
social_media_data$age_group <- cut(social_media_data$Age,
                                   breaks = breaks,
                                   labels = labels,
                                   right = F)
```

```{r}
social_media_data %>%
  select(age_group,`Video Category`) %>%
  table()
```

Entertainment and vlogs are the most popular video categories among users under the age of 21, while Jokes/Memes and Life Hacks are the two most popular categories for people 50 and older.

### 12

```{r}
social_media_data %>%
  select(age_group, `Platform`) %>%
  table()
```

TikTok and YouTube are the most popular platforms for both people under 21 and people over 50.

#### Additional Questions for 231 Students

13. Explore the data. What are some patterns that you notice? Create one to two visualizations.
14. Summarize your findings in a brief (1-2 paragraphs) report.
